%% Cover letter for the Omega submission — carries the author identity;
%% the manuscript itself is anonymised (\anontrue in main.tex).
%% Build: pdflatex cover_letter
\documentclass[11pt]{article}
\usepackage[a4paper,margin=2.6cm]{geometry}
\usepackage{xcolor}
\usepackage[colorlinks=true,allcolors=blue!50!black]{hyperref}
\usepackage{parskip}
\pagestyle{empty}

\begin{document}

\begin{flushright}
Quang-Vinh Dang\\
British University Vietnam\\
Hung Yen, Vietnam\\
\href{mailto:vinh.dq4@buv.edu.vn}{vinh.dq4@buv.edu.vn}\\[4pt]
\today
\end{flushright}

\vspace{6pt}
\noindent
The Editors-in-Chief\\
\emph{Omega --- The International Journal of Management Science}

\vspace{14pt}
\noindent Dear Editors,

Please consider the enclosed manuscript, \emph{``BATON:
optimal-stopping recourse for vehicle routing with simultaneous
pickup, delivery, and stochastic demands,''} for publication in
\emph{Omega}.

The paper addresses an operational question that sits squarely in the
journal's tradition of decision-making under uncertainty: when a
delivery vehicle that also collects pickups is filling up mid-route,
at which stop should the operator intervene---hand the remainder of
the route to a standby vehicle, detour to the depot to unload, or
keep driving? We show that this execution-stage question is a
finite-horizon optimal-stopping problem and solve it with a policy,
\textsc{Baton}, that prices every recourse action from the route's
real geometry and intervenes exactly when continuing becomes dearer
than the cheapest lever. The policy has no tunable threshold; we prove
that no fixed threshold can reproduce the optimal intervention
boundary, and that the endpoint-based risk labels used by prior work
are structurally blind to mid-route load peaks.

The computational study is, to our knowledge, the broadest comparison
of execution-stage recourse policies in this problem class:
eleven policies---including published rule-based restocking triggers,
rollout, depot-return restocking, a plug-in dynamic program, and a
GPU-trained reinforcement-learning baseline---evaluated on identical
plans from six planning regimes (three of them published robust
planners), on classical benchmarks and on new instances built from
OpenStreetMap road networks with customers at real retail locations in
five cities. All underlying routing plans are certified against
mixed-integer programming bounds, and every table and figure in the
paper regenerates mechanically from the result files.

This manuscript is original, has not been published previously, and is
not under consideration elsewhere. The authors---Quang-Vinh Dang and
Hoang-Viet Vu (British University Vietnam) and Phuc-Son Nguyen (UEH
University, Ho Chi Minh City)---have no conflicts of interest to
declare, and the work received no external funding; a separate
declaration of competing interests and a CRediT authorship statement
accompany this submission. The manuscript has been anonymised for
review, with author identity confined to this letter and the separate
title page. The code and data repository will be shared with the
editorial office on request during review and made public upon
acceptance.

Thank you for your consideration.

\vspace{14pt}
\noindent Sincerely, on behalf of all authors,

\vspace{6pt}
\noindent Quang-Vinh Dang (corresponding author)\\
British University Vietnam
\end{document}
